Numerical simulation of the Cordilleran ice sheet \\ through the last glacial cycle

### 2014-06-26 kappa.tex
Add parameter tables.

% My commands
%\newcommand{\note}[1]{\textbf{[NOTE: #1]}}
\newcommand{\todo}[1]{\emph{[\textbf{Todo:} #1]}}
%\newcommand{\aref}[0]{\textbf{[ref.]}}

@@

% Common notations
\newcommand{\doteps}[0]{\dot{\epsilon}} % epsilon dot
\newcommand{\IDT}[0]{\tens{\delta}}     % Identity tensor
\newcommand{\CST}[0]{\tens{\sigma}}     % Cauchy stress tensor
\newcommand{\DST}[0]{\tens{\tau}}       % deviatoric stress tensor
\newcommand{\SRT}[0]{\tens{\doteps}}    % strain-rate tensor
\newcommand{\vv}[0]{\vect{v}}           % velocity vector
\newcommand{\vsia}[0]{\vv_{\mathrm{SIA}}}   % SIA velocity
\newcommand{\vssa}[0]{\vv_{\mathrm{SSA}}}   % SSA velocity

% Common units
\newcommand{\unit}[1]{\ensuremath{\mathrm{#1}}}
\newcommand{\degree}[0]{\ensuremath{^{\circ}}}
\newcommand{\degC}[0]{\unit{{\degree}C}}

@@

\todo{Table. model variables and parameters}

@@

Thirdly, after defining the strain-rate tensor
${\SRT = \frac{1}{2}(\tgrad{\vv} + \tgrad{\vv}^{\mathrm{T}})}$,
and the deviatoric stress tensor, ${\DST = \CST - p\,\IDT}$,
where $p=\frac{1}{3}\tr(\CST)$ is the hydrostatic pressure and
$\tens{\delta}$ is the identity tensor, a \emph{constitutive law} for ice based
on laboratory experiments is used to relate these two quantities. It reads:
\begin{equation}
    \label{eqn:glenslaw}
    \SRT = A\,\tau_e^{n-1}\,\DST \,.
\end{equation}
The equivalent stress, $\tau_e$, is defined by
${\tau_e}^2 = -\mathrm{II}_{\DST} = \frac{1}{2} \tr(\DST^2)$,
where $\mathrm{II}_{\DST}$ is the second invariant of the stress tensor.
The ice softness coefficient, $A$, depends on ice temperature, pressure, and
water, $\omega$ through an Arrhenius-type law,
\begin{equation}
    A =
    \begin{cases}
        A_c (1+f\omega)\,e^\frac{-Q_c}{RT_{pa}}
            & \text{if}\ T < T_{crit} \,, \\
        A_w (1+f\omega)\,e^\frac{-Q_w}{RT_{pa}}
            & \text{if}\ T \ge T_{crit} \,,
    \end{cases}
\end{equation}
where $T_{pa}$ is the pressure-adjusted temperature calculated using the
Clapeyron relation, ${T_{pa} = T - \beta p}$.

@@

@@

@@

@@

\newcommand{\PDD}[0]{\mathrm{PDD}}
\newcommand{\sPDD}[0]{\sigma_{\mathrm{PDD}}}

At the interface between the ice sheet and the atmosphere (Fig.~??),
surfice mass fluxes are computed from monthly mean surface air temperature,
$T_m$, and monthly precipitation, $P_m$, by a temperature-index model
\citep[e.g.,][]{Hock.2003}. Surface accumulation, $a_s$, equals precipitation
when temperature is below ${T_0=0\,\degC}$, and decreases to zero linearly
with temperature between $T_0$ and ${T_1=2\,\degC}$,
\begin{equation}
    a_s =
    \begin{cases}
        0       & \text{if}\ T_m \le T_0 \,, \\
        P_m \frac{T_m-T_0}{T_1-T_0}
                & \text{if}\ T_0 < b < T_1 \,, \\
        P_m     & \text{if}\ T_1 \le T_m \,,
    \end{cases}
\end{equation}

@@

>>>>

% My commands
%\newcommand{\note}[1]{\textbf{[NOTE: #1]}}
\newcommand{\todo}[1]{\emph{[\textbf{Todo:} #1]}}
\newcommand{\mref}[0]{\textbf{[ref.]}}

@@

% Common notations
\newcommand{\doteps}[0]{\dot{\epsilon}} % epsilon dot
\newcommand{\IDT}[0]{\tens{\delta}}     % Identity tensor
\newcommand{\CST}[0]{\tens{\sigma}}     % Cauchy stress tensor
\newcommand{\DST}[0]{\tens{\tau}}       % deviatoric stress tensor
\newcommand{\SRT}[0]{\tens{\doteps}}    % strain-rate tensor
\newcommand{\vv}[0]{\vect{v}}           % velocity vector
\newcommand{\vsia}[0]{\vv_{\mathrm{SIA}}}   % SIA velocity
\newcommand{\vssa}[0]{\vv_{\mathrm{SSA}}}   % SSA velocity
\newcommand{\PDD}[0]{\mathrm{PDD}}
\newcommand{\sPDD}[0]{\sigma_{\mathrm{PDD}}}

% Common units
\newcommand{\e}[1]{\ensuremath{\times 10^{#1}}}
\newcommand{\unit}[1]{\ensuremath{\mathrm{#1}}}
\newcommand{\degree}[0]{\ensuremath{^{\circ}}}
\newcommand{\degC}[0]{\unit{{\degree}C}}

@@

@@

Thirdly, after defining the strain-rate tensor
${\SRT = \frac{1}{2}(\tgrad{\vv} + \tgrad{\vv}^{\mathrm{T}})}$,
and the deviatoric stress tensor, ${\DST = \CST - p\,\IDT}$,
where $p=\frac{1}{3}\tr(\CST)$ is the hydrostatic pressure and
$\tens{\delta}$ is the identity tensor, a \emph{constitutive law} for ice based
on laboratory experiments is used to relate these two quantities. It reads:
\begin{equation}
    \label{eqn:glenslaw}
    \SRT = A\,\tau_e^{n-1}\,\DST \,.
\end{equation}
The equivalent stress, $\tau_e$, is defined by
${\tau_e}^2 = -\mathrm{II}_{\DST} = \frac{1}{2} \tr(\DST^2)$,
where $\mathrm{II}_{\DST}$ is the second invariant of the stress tensor.
The ice softness coefficient, $A$, depends on ice temperature, pressure, and
water content through an Arrhenius-type law,
\begin{equation}
    A =
    \begin{cases}
        A_c (1+f\omega)\,e^\frac{-Q_c}{RT_{pa}}
            & \text{if}\ T < T_c \,, \\
        A_w (1+f\omega)\,e^\frac{-Q_w}{RT_{pa}}
            & \text{if}\ T \ge T_c \,,
    \end{cases}
\end{equation}
where $T_{pa}$ is the pressure-adjusted temperature calculated using the
Clapeyron relation, ${T_{pa} = T - \beta p}$. The water fraction, $\omega$, is
capped at a threshold of 0.01, above which no measurements are available.

@@

\begin{table}
  \centering
  \caption{Constant parameters of the ice flow model.
           \todo{find a better reference than \citet{Aschwanden.etal.2012}
                 for all the physical constants. Double-check
                 \citet{Paterson.Budd.1982} and add missing references.}}
  \label{tab:iceflowparams}
  \makebox[0pt]{\begin{tabular*}{165mm}{@{\extracolsep{\fill}}rlrll}
    \hline
    Not.    & Name & Value & Unit & Source \\
    \hline

    $\rho$  & Ice density
            & 910
            & \unit{kg\,m^{-3}}
            & \citet{Aschwanden.etal.2012} \\

    $g$     & Standard gravity
            & 9.81
            & \unit{m\,s^{-2}}
            & \citet{Aschwanden.etal.2012} \\

    $n$     & Glen exponent
            & 3
            & -
            & \mref \\

    $A_c$   & Ice hardness coefficient cold
            & 3.61\e{-13}
            & \unit{Pa^{-3}\,s^{-1}}
            & \citet{Paterson.Budd.1982} \\

    $A_w$   & Ice hardness coefficient warm
            & 1.73\e3
            & \unit{Pa^{-3}\,s^{-1}}
            & \citet{Paterson.Budd.1982} \\

    $Q_c$   & Flow law activation energy cold
            & 6.0\e4
            & \unit{J\,mol^{-1}}
            & \citet{Paterson.Budd.1982} \\

    $Q_w$   & Flow law activation energy warm
            & 13.9\e4
            & \unit{J\,mol^{-1}}
            & \citet{Paterson.Budd.1982} \\

    $R$     & Ideal gas constant
            & 8.31441
            & \unit{J\,mol^{-1}\,K^{-1}}
            & \mref \\

    $T_c$   & Flow law critical temperature
            & 263.15
            & \unit{K}
            & \citet{Paterson.Budd.1982} \\

    $f$     & Flow law water fraction coeff.
            & 181.25
            & -
            & \citet{Lliboutry.Duval.1985} \\

    $\beta$ & Clapeyron constant
            & 7.9\e{-8}
            & \unit{K\,Pa^{-1}}
            & \citet{Luthi.etal.2002} \\

    $c_i$   & Ice specific heat capacity
            & 2009
            & \unit{J\,kg^{-1}\,K^{-1}}
            & \citet{Aschwanden.etal.2012} \\

    $c_w$   & Water specific heat capacity
            & 4170
            & \unit{J\,kg^{-1}\,K^{-1}}
            & \citet{Aschwanden.etal.2012} \\

    $k$     & Ice thermal conductivity
            & 2.10
            & \unit{J\,m^{-1}\,K^{-1}\,s^{-1}}
            & \citet{Aschwanden.etal.2012} \\

    $L$     & Water latent heat of fusion
            & 3.34\e5
            & \unit{J\,kg^{-1}\,K^{-1}}
            & \citet{Aschwanden.etal.2012} \\

    \hline
  \end{tabular*}}
\end{table}

@@

\begin{table}
  \centering
  \caption{Constant parameters of the sliding model. Missing sources
           correspond to uncalibrated values.}
  \label{tab:slidingparams}
  \makebox[0pt]{\begin{tabular*}{165mm}{@{\extracolsep{\fill}}rlrll}
    \hline
    Not.    & Name & Value & Unit & Source \\
    \hline

    $q$     & Pseudo-plastic sliding exponent
            & 0.25
            & -
            & - \\

    $v_{th}$& Pseudo-plastic threshold velocity
            & 100.0
            & \unit{m\,yr^{-1}}
            & - \\

    $c_0$   & Till cohesion
            & 0.0
            & Pa
            & - \\

    $\alpha$& \todo{check value}
            & 0.95
            & -
            & - \\

    $\delta$& \todo{find a good name}
            & 0.02
            & -
            & - \\

    $e_0$   & Till reference void ratio
            & 0.69
            & -
            & \citet{Tulaczyk.etal.2000} \\

    $C_c$   & Till compressibility coefficient
            & 0.12
            & -
            & \citet{Tulaczyk.etal.2000} \\

    $W_{max}$ & Maximal till water thickness
            & 2.0
            & m
            & - \\

    $b_0$   & Altitude of max. friction angle
            & 0
            & m
            & - \\

    $b_1$   & Altitude of min. friction angle
            & 200
            & m
            & - \\

    $\phi_0$& Minimum friction angle
            & 10/15
            & \degree
            & - \\

    $\phi_1$& Maximum friction angle
            & 30/45
            & \degree
            & - \\

    \hline
  \end{tabular*}}
\end{table}

@@

\begin{table}
  \centering
  \caption{Constant parameters of the bedrock model.
           \todo{Find out what the bedrock density corresponds to. Track down
                 original sources in \citet{Ritz.1997}}}
  \label{tab:bedrockparams}
  \makebox[0pt]{\begin{tabular*}{165mm}{@{\extracolsep{\fill}}rlrll}
    \hline
    Not.    & Name & Value & Unit & Source \\
    \hline

    $\rho_b$& Bedrock density
            & 3300
            & \unit{kg\,m^{-3}}
            & \mref \\

    $c_b$   & Bedrock specific heat capacity
            & 1000
            & \unit{J\,kg^{-1}\,K^{-1}}
            & \citet{Ritz.1997} \\

    $k_b$   & Bedrock thermal conductivity
            & 3.0
            & \unit{J\,m^{-1}\,K^{-1}\,s^{-1}}
            & \citet{Ritz.1997} \\

    $\nu_m$ & Mantle viscosity
            & 1\e21
            & \unit{Pa\,s}
            & \citet{Lingle.Clark.1985} \\

    $\rho_l$& Lithosphere density
            & 3300
            & \unit{kg\,m^{-3}}
            & \citet{Lingle.Clark.1985} \\

    $D$     & Lithosphere flexural rigidity
            & 5.0\e24
            & \unit{N}
            & \citet{Lingle.Clark.1985} \\

    \hline
%	pism overrides:bootstrapping geothermal flux value no var = 0.070 ;
  \end{tabular*}}
\end{table}

@@

At the interface between the ice sheet and the atmosphere (Fig.~??),
surfice mass fluxes are computed from monthly mean surface air temperature,
$T_m$, and monthly precipitation, $P_m$, by a temperature-index model
\citep[e.g.,][]{Hock.2003}. Surface accumulation, $a_s$, equals precipitation
when temperature is below ${T_0=0\,\degC}$, and decreases to zero linearly
with temperature between $T_0$ and ${T_1=2\,\degC}$,
\begin{equation}
    a_s =
    \begin{cases}
        0       & \text{if}\ T_m \le T_s \,, \\
        P_m \frac{T_m-T_s}{T_r-T_s}
                & \text{if}\ T_s < T_m < T_r \,, \\
        P_m     & \text{if}\ T_r \le T_m \,,
    \end{cases}
\end{equation}

@@

\begin{table}
  \centering
  \caption{Constant parameters of the surface and atmosphere models.
           \todo{Find a reference for the snow and rain temperature thresholds,
                 maybe \citet{Hock.2003}.}}
  \label{tab:surfaceparams}
  \makebox[0pt]{\begin{tabular*}{165mm}{@{\extracolsep{\fill}}rlrll}
    \hline
    Not.    & Name & Value & Unit & Source \\
    \hline

    $T_s$   & Temperature of snow precipitation
            & 273.15
            & \unit{K}
            & \mref \\

    $T_r$   & Temperature of rain precipitation
            & 275.15
            & \unit{K}
            & \mref \\

    $F_s$   & Degree-day factor for ice
            & 3.04\e{-3}
            & \unit{m\,K^{-1}\,day^{-1}}
            & \citet{Shea.etal.2009} \\

    $F_i$   & Degree-day factor for ice
            & 4.59\e{-3}
            & \unit{m\,K^{-1}\,day^{-1}}
            & \citet{Shea.etal.2009} \\

    $\gamma$& Air temperature lapse-rate
            & 6\e{-3}
            & \unit{K\,m{-1}}
            & - \\

    \hline
  \end{tabular*}}
\end{table}
